Extending the Advanced Forensic Format to accommodate Multiple Data Sources, Logical Evidence and Forensic Workflow

### 2009-03-10 paper/paper.tex
simson1: Ignore-this: cd3848f5f55f9e0c9da0835e6bc0d37c
Initial work with the new section describing changes to AFF.

@@

%\begin{frontmatter}
\begin{document}
\title{The Forensic Interchange Format - a file format for digital evidence}
\author{M. I. Cohen\footnote{scudette@gmail.com} and Bradley Schatz}
%\author{M. I. Cohen}
%\address{M. I. Cohen is a Data specialist with the Australian Federal Police, Brisbane, Australia}
%\baselineskip=20pt %Sets line spacing to 1 unit
%\end{frontmatter}
%\thanks{M. I. Cohen is a Data specialist with the Australian Federal Police, Brisbane, Australia}%
\maketitle

\begin{abstract}
Forensics analysis requires the acquisition and management of many
different types of evidence material. Often the same evidence is
reviewed by several different tools. We propose an open, extensible
file format for storing of evidence and interchanging analysis results
among different tools. The Forensic Interchange Format (FIF) was
designed to be simple to implement, relying on the well supported Zip
specifications for bit level file access.
\end{abstract}

@@

>>>>

% DFRWS Call Information:
% http://www.dfrws.org/2009/cfp.shtml
% Important Dates
% Submission deadline: March 16, 2009 (any time zone)
% Author notification: April 28, 2009
% Final draft due: May 19, 2009
% Pre-conference Workshops: August 16, 2009
% Conference dates: August 17-19, 2009

% Publication Criteria
% Research papers must be original contributions, not substantially
% duplicate previous work, and must not be under simultaneous
% publication review elsewhere. The review process will be
% ``double-blind'' (the reviewers will not know who the authors are, and
% the authors will not know who the reviewers are). Therefore, the
% version submitted for review should not contain the names or
% affiliations of the authors. When referring to their own previous
% work, authors should use the third person instead of the first person
% (i.e. ``Smith and Jones [2] previously determined...'' instead of ``We
% [2] previously determined..''). Authors are expected to present their
% work in person at the workshop and must have at least one registration
% per paper in order to be included in the proceedings.

@@

%\begin{frontmatter}
\begin{document}
\title{Extending the Advanced Forensic Format to accommodate Multiple
  Data Sources, Logical Evidence and Forensic Workflow}
\author{Blinded for review}
%\author{M. I. Cohen\footnote{scudette@gmail.com}, Simson Garfinkel, and Bradley Schatz}
%\author{M. I. Cohen}
%\address{M. I. Cohen is a Data specialist with the Australian Federal Police, Brisbane, Australia}
%\baselineskip=20pt %Sets line spacing to 1 unit
%\end{frontmatter}
%\thanks{M. I. Cohen is a Data specialist with the Australian Federal Police, Brisbane, Australia}
\maketitle

\begin{abstract}
Forensics analysis requires the acquisition and management of many
different types of evidence material, including individual disk
drives, RAID sets, network packets, memory images, and extracted files. Often the same evidence is
reviewed by several different tools or examiners in different
locations. We propose a backwards-compatible evolutionary redesign of the Advanced
Forensic Format an open, extensible
file format for storing of evidence and interchanging analysis results
among different tools. The Forensic Interchange Format (FIF) was
designed to be simple to implement, relying on the well supported Zip
specifications for bit level file access.
\end{abstract}

@@

Unlike other kinds of evidence, digital evidence can be instantly lost
or corrupted if technical measures are not immediately taken to
preserve its existence and integrity. As a result, preserving digital
evidence is an important part of most digital
investigations\cite{carrier:event-based}.

In recent years there has been a steady and growing interest in the
actual file formats and containers used to store this digital
evidence. Early practitioners created exact bit-for-bit copies (``mirror disks'') employed
proprietary software systems (e.g. \cite{safeback,ilook,encase}) for
making and authenticating ``images'' of digital evidence. PyFlag\cite{pyflag}
introduced a ``seekable gzip'' format that allowed disk images to be
stored in a form that was compressed but allowed the random-access to
evidence data necessary for forensic analysis. The Advanced Forensic
Format (AFF) expanded on this idea with a forensic file format that
allowed both data and arbitrary metadata to be stored in a single
digital archive\cite{garfinkel:aff}. Recently the open source AFF
implementation library (AFFLIB) was expanded to support both
encryption and digital signatures of digital evidence\cite{garfinkel:affcrypto}.

This paper extends the previous work by adapting AFF to the multiple
heterogeneous data types that might arise in a modern digital
investigation, including data from multiple data storage devices, new
data types (including network packets and memory images), extracted
logical evidence, and forensic workflow. We do this in a manner that
is at once upwards compatible with AFF while adopting the AFF
bit-level specification so that AFF files can be processed using
standard non-forensic tools.

\section{A New Forensic Format}

AFF's flexibility comes from the fact that forensic data and metadata
is stored as arbitrary name/value pairs called \emph{segments}. For example, the first 16MB
of a disk image is stored in a segment called \texttt{page0}, the
second 16MB in a segment called \texttt{page1}, \emph{etc.} Because of
this flexibility, it was
relatively easy for Garfinkel to extend AFF to support encryption,
digital signatures, and the storage of new kinds of metadata such as
chain-of-custody information\cite{garfinkel:affcrypto}. .

\subsection{AFF Limitations}
After imaging literally thousands of devices using Garfinkel's AFF
tools and adding support for AFF to numerous open source programs, we
observed numerous problems in the underlying standard
and Garfinkel's AFFLIB implementation:

\begin{itemize}
\item While AFF's design stores a single disk image in each evidence
  file, modern digital investigations typically involve many seized
  computers or pieces of media. 
\item AFF has no provision for storing memory images or intercepted
  network packets.
\item AFF has no provisions for storing extracted files that is
  analogous to the EnCase ``Logical Evidence File'' (L01) format.
\item AFF's encryption system leaks information about the contents of
  an evidence file because segment names are not encrypted.
\item AFF's default compression page size of 16MB can impose significant overhead
  when accessing NTFS Master File Tables (MFT), as these structures
  tend to be highly fragmented on systems that have seen significant
  use. 
\item Although the AFF specification calls for a ``table of contents'' similar
  to the ZIP\cite{zip-format} ``central directory'' that is stored at the end of AFF
  files, Garfinkel never implemented this directory in the publicly
  released AFFLIB. As a result, every header of every segment in an
  AFF file needs to be read when a file is opened. It practice this
  can take 10-30 seconds the first time a large AFF file is
  opened. (Once the file is opened, the sectors corresponding to the
  segment headers are stored in the computer's disk cache.)
\item AFF's bit-level specification is essentially a simple container
  file specification. Given the that there are other container file
  specifications that are much more widely supported with both
  developer and end-user tools, it seemed reasonable to migrate AFF
  from its home-grown format to one of the existing standards.
\end{itemize}

We have addressed these issues with a three step process. First, we
have extended the AFF schema to support multiple forensic images
(which are now called \emph{streams}) in a single evidence
file. Second, we have changed the on disk format, jettisoning AFF1.0's
unique container format and replacing it with the ZIP64 file
format. Third, we are adopting AFFLIB to implement the new schema and
file format, which will allow existing AFF-aware programs to access
files written in the new format without modification. We call this new
format AFF2.

\subsection{AFF2: New Features}
AFF2 is designed to include the following functionality:

\begin{itemize}
\item Evidence is stored in \emph{volumes}. A volume can consist of a
  single file (\emph{e.g.} a file with the \texttt{.aff} extension), multiple files (\texttt{.afd}), an
  object storage system (such as Amazon S3), or a hybrid (for example,
  metadata stored in an SQL database with physical data stored in
  large files in a file system.)
\item A single volume can contain evidence from one or more physical
  devices.
\item RAID information can be directly encoded, allowing data within a
  RAID to be logically constructed at read time.
\item Although the page size remains 16MB by default, each 16MB page
  can be built from multiple compressed chunks (8K size by default),
  significantly decreasing the overhead to access a few blocks inside
  a compressed page.
\item The bit-level container file format has changed from AFFs's
  original container file format to ZIP64, allowing improved
  integration with existing developer tools.
\item Transparent upwards compatibility, so that programs linked with
  the AFFLIB implementation or the pyaff.py python library can
  interchangeably read either AFF1 or AFF2 files. 
\end{itemize}
